---
title: "Untitled"
output: html_document
date: "2024-11-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# EDA

# FLIR (Forward-Looking Infrared) Groups:

The dataset includes data captured using the FLIR thermography camera:
 
 - FLIR Group 1: Ambient Temp 20-24°C, Relative Humidity: 10–62%
 
 - FLIR Group 2: Ambient Temp 24-29°C, Relative Humidity: 10–62%
 
 
### Load and clean FLIR GROUP 1
```{r}
flir_group1 <- read.csv("FLIR_group1.csv", header = FALSE, skip = 2)
colnames(flir_group1) <- flir_group1[1, ]
flir_group1 <- flir_group1[-1, ]

# Replace empty strings or spaces with NA
flir_group1[flir_group1 == "" | flir_group1 == " "] <- NA

# Remove columns where all values are NA
flir_group1_clean <- flir_group1[, colSums(is.na(flir_group1)) != nrow(flir_group1)]
```


### Load and clean FLIR GROUP 2
```{r}
flir_group2 <- read.csv("FLIR_group2.csv", header = FALSE, skip = 2)
colnames(flir_group2) <- flir_group2[1, ]
flir_group2 <- flir_group2[-1, ]

# Replace empty strings or spaces with NA
flir_group2[flir_group2 == "" | flir_group2 == " "] <- NA

# Remove columns where all values are NA
flir_group2_clean <- flir_group2[, colSums(is.na(flir_group2)) != nrow(flir_group2)]
```





# ICI (Infrared Cameras Inc.) Groups:

- ICI Group 1: Ambient Temp 20-24°C

- ICI Group 2: Ambient Temp 24-29°C

### Load and clean ICI Group 1
```{r}
ici_group1 <- read.csv("ICI_group1.csv", header = FALSE, skip = 2)
colnames(ici_group1) <- ici_group1[1, ]
ici_group1 <- ici_group1[-1, ]
```


### Replace empty strings or spaces with NA
```{r}
ici_group1[ici_group1 == "" | ici_group1 == " "] <- NA
```


### Remove columns where all values are NA
```{r}
ici_group1_clean <- ici_group1[, colSums(is.na(ici_group1)) != nrow(ici_group1)]
```


### Load and clean ICI Group 2
```{r}
ici_group2 <- read.csv("ICI_group2.csv", header = FALSE, skip = 2)
colnames(ici_group2) <- ici_group2[1, ]
ici_group2 <- ici_group2[-1, ]
```


### Replace empty strings or spaces with NA
```{r}
ici_group2[ici_group2 == "" | ici_group2 == " "] <- NA
```


### Remove columns where all values are NA
```{r}
ici_group2_clean <- ici_group2[, colSums(is.na(ici_group2)) != nrow(ici_group2)]
```




# Check the structure and column names of the cleaned datasets
```{r}
str(flir_group1_clean)
str(flir_group2_clean)
str(ici_group1_clean)
str(ici_group2_clean)
```

```{r}
names(flir_group1_clean)
names(flir_group2_clean)
names(ici_group1_clean)
names(ici_group2_clean)
```

# Check for duplicate SubjectIDs and overlaps
Ensure all SubjectIDs are unique within each dataset
```{r}
length(unique(flir_group1_clean$SubjectID)) == nrow(flir_group1_clean)
length(unique(flir_group2_clean$SubjectID)) == nrow(flir_group2_clean)
length(unique(ici_group1_clean$SubjectID)) == nrow(ici_group1_clean)
length(unique(ici_group2_clean$SubjectID)) == nrow(ici_group2_clean)
```


# Find duplicated SubjectIDs
```{r}
duplicated_flir_group1 <- flir_group1_clean$SubjectID[duplicated(flir_group1_clean$SubjectID)]
duplicated_flir_group2 <- flir_group2_clean$SubjectID[duplicated(flir_group2_clean$SubjectID)]
duplicated_ici_group1 <- ici_group1_clean$SubjectID[duplicated(ici_group1_clean$SubjectID)]
duplicated_ici_group2 <- ici_group2_clean$SubjectID[duplicated(ici_group2_clean$SubjectID)]
```

# Check for overlapping SubjectIDs between datasets
```{r}
overlap_flir1_flir2 <- intersect(flir_group1_clean$SubjectID, flir_group2_clean$SubjectID)
overlap_flir1_ici1 <- intersect(flir_group1_clean$SubjectID, ici_group1_clean$SubjectID)
overlap_flir1_ici2 <- intersect(flir_group1_clean$SubjectID, ici_group2_clean$SubjectID)
overlap_flir2_ici1 <- intersect(flir_group2_clean$SubjectID, ici_group1_clean$SubjectID)
overlap_flir2_ici2 <- intersect(flir_group2_clean$SubjectID, ici_group2_clean$SubjectID)
overlap_ici1_ici2 <- intersect(ici_group1_clean$SubjectID, ici_group2_clean$SubjectID)
```

# Print overlaps
```{r}
overlap_flir1_flir2
overlap_flir1_ici1
overlap_flir1_ici2
overlap_flir2_ici1
overlap_flir2_ici2
overlap_ici1_ici2
```





# Clean column names and convert variables

### Fix incorrect variable name "T_LC14" to "T_LC4" in all datasets
```{r}
fix_column_name <- function(data) {
  if ("T_LC14" %in% colnames(data)) {
    colnames(data)[colnames(data) == "T_LC14"] <- "T_LC4"
  }
  return(data)
}
```


### Apply the function to all datasets
```{r}
flir_group1_clean <- fix_column_name(flir_group1_clean)
flir_group2_clean <- fix_column_name(flir_group2_clean)
ici_group1_clean <- fix_column_name(ici_group1_clean)
ici_group2_clean <- fix_column_name(ici_group2_clean)
```


# Vector of base variable names excluding the round numbers
```{r}
base_temp_vars <- c("T_offset", "Max1R13_", "Max1L13_", "aveAllR13_", "aveAllL13_", 
                    "T_RC", "T_RC_Dry", "T_RC_Wet", "T_RC_Max", 
                    "T_LC", "T_LC_Dry", "T_LC_Wet", "T_LC_Max", 
                    "RCC", "LCC", "canthiMax", "canthi4Max", 
                    "T_FHCC", "T_FHRC", "T_FHLC", "T_FHBC", "T_FHTC", 
                    "T_FH_Max", "T_FHC_Max", "T_Max", "T_OR", "T_OR_Max")
```


# Function to convert variables to numeric for all 4 rounds
```{r}
convert_to_numeric <- function(data, base_vars, rounds = 4) {
  for (round in 1:rounds) {
    for (var in base_vars) {
      var_name <- paste0(var, round)
      if (var_name %in% colnames(data)) {
        data[[var_name]] <- as.numeric(data[[var_name]])
      } else {
        cat(var_name, "not found in the dataset\n")
      }
    }
  }
  return(data)
}
```


# Apply the function to all datasets
```{r}
flir_group1_clean <- convert_to_numeric(flir_group1_clean, base_temp_vars)
flir_group2_clean <- convert_to_numeric(flir_group2_clean, base_temp_vars)
ici_group1_clean <- convert_to_numeric(ici_group1_clean, base_temp_vars)
ici_group2_clean <- convert_to_numeric(ici_group2_clean, base_temp_vars)
```


# Confirm the conversion
```{r}
str(flir_group1_clean)
str(flir_group2_clean)
str(ici_group1_clean)
str(ici_group2_clean)
```



# Convert other specific columns to numeric or factors

```{r}
# Convert aveOralF and aveOralM to numeric
convert_oral_vars <- function(data) {
  data$aveOralF <- as.numeric(data$aveOralF)
  data$aveOralM <- as.numeric(data$aveOralM)
  return(data)
}

flir_group1_clean <- convert_oral_vars(flir_group1_clean)
flir_group2_clean <- convert_oral_vars(flir_group2_clean)
ici_group1_clean <- convert_oral_vars(ici_group1_clean)
ici_group2_clean <- convert_oral_vars(ici_group2_clean)
```


# Convert Gender, Age, Ethnicity and Cosmetics to factors
```{r}
convert_factors <- function(data) {
  data$Gender <- as.factor(data$Gender)
  data$Age <- as.factor(data$Age)
  data$Ethnicity <- as.factor(data$Ethnicity)
  data$Cosmetics <- as.factor(data$Cosmetics)
  return(data)
}

flir_group1_clean <- convert_factors(flir_group1_clean)
flir_group2_clean <- convert_factors(flir_group2_clean)
ici_group1_clean <- convert_factors(ici_group1_clean)
ici_group2_clean <- convert_factors(ici_group2_clean)
```


# Convert T_atm, Humidity, and Distance to numeric
```{r}
convert_environment_vars <- function(data) {
  data$T_atm <- as.numeric(data$T_atm)
  data$Humidity <- as.numeric(data$Humidity)
  data$Distance <- as.numeric(data$Distance)
  return(data)
}

flir_group1_clean <- convert_environment_vars(flir_group1_clean)
flir_group2_clean <- convert_environment_vars(flir_group2_clean)
ici_group1_clean <- convert_environment_vars(ici_group1_clean)
ici_group2_clean <- convert_environment_vars(ici_group2_clean)
```




# Convert Date and Time columns
```{r}
# Use lubridate to convert Date and Time
library(lubridate)

convert_date_time <- function(data) {
  data$Time <- lubridate::hms(data$Time)   # Convert Time to HH:MM:SS format
  data$Date <- lubridate::ymd(data$Date)   # Convert Date to YYYY-MM-DD format
  return(data)
}
```


# Apply the function to all datasets
```{r}
flir_group1_clean <- convert_date_time(flir_group1_clean)
flir_group2_clean <- convert_date_time(flir_group2_clean)
ici_group1_clean <- convert_date_time(ici_group1_clean)
ici_group2_clean <- convert_date_time(ici_group2_clean)
```


# Check the summaries to confirm correct formatting
```{r}
summary(flir_group1_clean$Time)
summary(flir_group1_clean$Date)

summary(flir_group2_clean$Time)
summary(flir_group2_clean$Date)

summary(ici_group1_clean$Time)
summary(ici_group1_clean$Date)

summary(ici_group2_clean$Time)
summary(ici_group2_clean$Date)
```




# Handling the NA values #
```{r}
library(naniar)
```


# Visualize missing data for each dataset
```{r}
vis_miss(flir_group1_clean)
vis_miss(flir_group2_clean)
vis_miss(ici_group1_clean)
vis_miss(ici_group2_clean)
```




# Function to check unique levels for categorical variables ####
```{r}

check_categorical_levels <- function(data) {
  for (column in colnames(data)) {
    
    # If the column is a factor, print its unique levels and number of NA values
    
    if (is.factor(data[[column]])) {
      cat("\nUnique values in", column, ":\n")
      print(unique(data[[column]]))
      cat("Number of NAs in", column, ":\n")
      print(sum(is.na(data[[column]])))
    }
  }
  return(data)
}
```


# Apply to all datasets

```{r}
# For FLIR Group 1
cat("FLIR Group 1:\n")
flir_group1_clean <- check_categorical_levels(flir_group1_clean)

# For FLIR Group 2
cat("\nFLIR Group 2:\n")
flir_group2_clean <- check_categorical_levels(flir_group2_clean)

# For ICI Group 1
cat("\nICI Group 1:\n")
ici_group1_clean <- check_categorical_levels(ici_group1_clean)

# For ICI Group 2
cat("\nICI Group 2:\n")
ici_group2_clean <- check_categorical_levels(ici_group2_clean)
```

# Removing rows with missing values 

```{r}
# complete.cases to remove rows that contain any missing values from each dataset
flir_group1_clean <- flir_group1_clean[complete.cases(flir_group1_clean), ]
flir_group2_clean <- flir_group2_clean[complete.cases(flir_group2_clean), ]
ici_group1_clean <- ici_group1_clean[complete.cases(ici_group1_clean), ]
ici_group2_clean <- ici_group2_clean[complete.cases(ici_group2_clean), ]
```


# Count the number of remaining NA values
```{r}
cat("\nRemaining NA counts:\n")
cat("FLIR Group 1: ", sum(is.na(flir_group1_clean)), "\n")
cat("FLIR Group 2: ", sum(is.na(flir_group2_clean)), "\n")
cat("ICI Group 1: ", sum(is.na(ici_group1_clean)), "\n")
cat("ICI Group 2: ", sum(is.na(ici_group2_clean)), "\n")
```


# Final structure check

## Check the structure of the cleaned datasets after removing rows with missing values
```{r}
cat("\nFinal structure of datasets:\n")
str(flir_group1_clean) # Expected: 509 rows, ~7% NAs before removal
str(flir_group2_clean) # Expected: 450 rows, ~5.5% NAs before removal
str(ici_group1_clean)  # Expected: 364 rows, ~32% NAs before removal
str(ici_group2_clean)  # Expected: 305 rows, ~35% NAs before removal
```

# Save cleaned datasets to CSV files

```{r}
# For flir_group1_clean
write.csv(flir_group1_clean, file = "/Users/karibautista/Desktop/flir_group1_clean.csv", row.names = FALSE)

# For flir_group2_clean
write.csv(flir_group2_clean, file = "/Users/karibautista/Desktop/flir_group2_clean.csv", row.names = FALSE)

# For ici_group1_clean
write.csv(ici_group1_clean, file = "/Users/karibautista/Desktop/ici_group1_clean.csv", row.names = FALSE)

# For ici_group2_clean
write.csv(ici_group2_clean, file = "/Users/karibautista/Desktop/ici_group2_clean.csv", row.names = FALSE)

cat("\nDatasets have been saved as CSV files on the Desktop.\n")
```

#-----------------------------------------------------------------------------------------------------------

# Create a new column with the average of the four rounds
## FLIR GROUP1 AND GROUP2
```{r}
df <- read.csv("flir_group1_clean.csv")
df2 <- read.csv("flir_group2_clean.csv")
```

GROUP 1
Calculate the average of the T_offset variables across the four rounds
```{r}
df$T_offset_avg <- rowMeans(df[, c("T_offset1", "T_offset2", "T_offset3", "T_offset4")], na.rm = TRUE)
df$Max1R_avg <- rowMeans(df[, c("Max1R13_1", "Max1R13_2", "Max1R13_3", "Max1R13_4")], na.rm = TRUE)
df$Max1L_avg <- rowMeans(df[, c("Max1L13_1", "Max1L13_2", "Max1L13_3", "Max1L13_4")], na.rm = TRUE)
df$aveAllR_avg <- rowMeans(df[, c("aveAllR13_1", "aveAllR13_2", "aveAllR13_3", "aveAllR13_4")], na.rm = TRUE)
df$aveAllL_avg <- rowMeans(df[, c("aveAllL13_1", "aveAllL13_2", "aveAllL13_3", "aveAllL13_4")], na.rm = TRUE)
df$T_RC_avg <- rowMeans(df[, c("T_RC1", "T_RC2", "T_RC3", "T_RC4")], na.rm = TRUE)
df$T_RC_Dry_avg <- rowMeans(df[, c("T_RC_Dry1", "T_RC_Dry2", "T_RC_Dry3", "T_RC_Dry4")], na.rm = TRUE)
df$T_RC_Wet_avg <- rowMeans(df[, c("T_RC_Wet1", "T_RC_Wet2", "T_RC_Wet3", "T_RC_Wet4")], na.rm = TRUE)
df$T_RC_Max_avg <- rowMeans(df[, c("T_RC_Max1", "T_RC_Max2", "T_RC_Max3", "T_RC_Max4")], na.rm = TRUE)
df$T_LC_avg <- rowMeans(df[, c("T_LC1", "T_LC2", "T_LC3", "T_LC4")], na.rm = TRUE)
df$T_LC_Dry_avg <- rowMeans(df[, c("T_LC_Dry1", "T_LC_Dry2", "T_LC_Dry3", "T_LC_Dry4")], na.rm = TRUE)
df$T_LC_Wet_avg <- rowMeans(df[, c("T_LC_Wet1", "T_LC_Wet2", "T_LC_Wet3", "T_LC_Wet4")], na.rm = TRUE)
df$T_LC_Max_avg <- rowMeans(df[, c("T_LC_Max1", "T_LC_Max2", "T_LC_Max3", "T_LC_Max4")], na.rm = TRUE)
df$RCC_avg <- rowMeans(df[, c("RCC1", "RCC2", "RCC3", "RCC4")], na.rm = TRUE)
df$LCC_avg <- rowMeans(df[, c("LCC1", "LCC2", "LCC3", "LCC4")], na.rm = TRUE)
df$canthiMax_avg <- rowMeans(df[, c("canthiMax1", "canthiMax2", "canthiMax3", "canthiMax4")], na.rm = TRUE)
df$canthi4Max_avg <- rowMeans(df[, c("canthi4Max1", "canthi4Max2", "canthi4Max3", "canthi4Max4")], na.rm = TRUE)
df$T_FHCC_avg <- rowMeans(df[, c("T_FHCC1", "T_FHCC2", "T_FHCC3", "T_FHCC4")], na.rm = TRUE)
df$T_FHRC_avg <- rowMeans(df[, c("T_FHRC1", "T_FHRC2", "T_FHRC3", "T_FHRC4")], na.rm = TRUE)
df$T_FHLC_avg <- rowMeans(df[, c("T_FHLC1", "T_FHLC2", "T_FHLC3", "T_FHLC4")], na.rm = TRUE)
df$T_FHBC_avg <- rowMeans(df[, c("T_FHBC1", "T_FHBC2", "T_FHBC3", "T_FHBC4")], na.rm = TRUE)
df$T_FHTC_avg <- rowMeans(df[, c("T_FHTC1", "T_FHTC2", "T_FHTC3", "T_FHTC4")], na.rm = TRUE)
df$T_FH_Max_avg <- rowMeans(df[, c("T_FH_Max1", "T_FH_Max2", "T_FH_Max3", "T_FH_Max4")], na.rm = TRUE)
df$T_FHC_Max_avg <- rowMeans(df[, c("T_FHC_Max1", "T_FHC_Max2", "T_FHC_Max3", "T_FHC_Max4")], na.rm = TRUE)
df$T_Max_avg <- rowMeans(df[, c("T_Max1", "T_Max2", "T_Max3", "T_Max4")], na.rm = TRUE)
df$T_OR_avg <- rowMeans(df[, c("T_OR1", "T_OR2", "T_OR3", "T_OR4")], na.rm = TRUE)
df$T_OR_Max_avg <- rowMeans(df[, c("T_OR_Max1", "T_OR_Max2", "T_OR_Max3", "T_OR_Max4")], na.rm = TRUE)
```

GROUP 2
```{r}
df2$T_offset_avg <- rowMeans(df2[, c("T_offset1", "T_offset2", "T_offset3", "T_offset4")], na.rm = TRUE)
df2$Max1R_avg <- rowMeans(df2[, c("Max1R13_1", "Max1R13_2", "Max1R13_3", "Max1R13_4")], na.rm = TRUE)
df2$Max1L_avg <- rowMeans(df2[, c("Max1L13_1", "Max1L13_2", "Max1L13_3", "Max1L13_4")], na.rm = TRUE)
df2$aveAllR_avg <- rowMeans(df2[, c("aveAllR13_1", "aveAllR13_2", "aveAllR13_3", "aveAllR13_4")], na.rm = TRUE)
df2$aveAllL_avg <- rowMeans(df2[, c("aveAllL13_1", "aveAllL13_2", "aveAllL13_3", "aveAllL13_4")], na.rm = TRUE)
df2$T_RC_avg <- rowMeans(df2[, c("T_RC1", "T_RC2", "T_RC3", "T_RC4")], na.rm = TRUE)
df2$T_RC_Dry_avg <- rowMeans(df2[, c("T_RC_Dry1", "T_RC_Dry2", "T_RC_Dry3", "T_RC_Dry4")], na.rm = TRUE)
df2$T_RC_Wet_avg <- rowMeans(df2[, c("T_RC_Wet1", "T_RC_Wet2", "T_RC_Wet3", "T_RC_Wet4")], na.rm = TRUE)
df2$T_RC_Max_avg <- rowMeans(df2[, c("T_RC_Max1", "T_RC_Max2", "T_RC_Max3", "T_RC_Max4")], na.rm = TRUE)
df2$T_LC_avg <- rowMeans(df2[, c("T_LC1", "T_LC2", "T_LC3", "T_LC4")], na.rm = TRUE)
df2$T_LC_Dry_avg <- rowMeans(df2[, c("T_LC_Dry1", "T_LC_Dry2", "T_LC_Dry3", "T_LC_Dry4")], na.rm = TRUE)
df2$T_LC_Wet_avg <- rowMeans(df2[, c("T_LC_Wet1", "T_LC_Wet2", "T_LC_Wet3", "T_LC_Wet4")], na.rm = TRUE)
df2$T_LC_Max_avg <- rowMeans(df2[, c("T_LC_Max1", "T_LC_Max2", "T_LC_Max3", "T_LC_Max4")], na.rm = TRUE)
df2$RCC_avg <- rowMeans(df2[, c("RCC1", "RCC2", "RCC3", "RCC4")], na.rm = TRUE)
df2$LCC_avg <- rowMeans(df2[, c("LCC1", "LCC2", "LCC3", "LCC4")], na.rm = TRUE)
df2$canthiMax_avg <- rowMeans(df2[, c("canthiMax1", "canthiMax2", "canthiMax3", "canthiMax4")], na.rm = TRUE)
df2$canthi4Max_avg <- rowMeans(df2[, c("canthi4Max1", "canthi4Max2", "canthi4Max3", "canthi4Max4")], na.rm = TRUE)
df2$T_FHCC_avg <- rowMeans(df2[, c("T_FHCC1", "T_FHCC2", "T_FHCC3", "T_FHCC4")], na.rm = TRUE)
df2$T_FHRC_avg <- rowMeans(df2[, c("T_FHRC1", "T_FHRC2", "T_FHRC3", "T_FHRC4")], na.rm = TRUE)
df2$T_FHLC_avg <- rowMeans(df2[, c("T_FHLC1", "T_FHLC2", "T_FHLC3", "T_FHLC4")], na.rm = TRUE)
df2$T_FHBC_avg <- rowMeans(df2[, c("T_FHBC1", "T_FHBC2", "T_FHBC3", "T_FHBC4")], na.rm = TRUE)
df2$T_FHTC_avg <- rowMeans(df2[, c("T_FHTC1", "T_FHTC2", "T_FHTC3", "T_FHTC4")], na.rm = TRUE)
df2$T_FH_Max_avg <- rowMeans(df2[, c("T_FH_Max1", "T_FH_Max2", "T_FH_Max3", "T_FH_Max4")], na.rm = TRUE)
df2$T_FHC_Max_avg <- rowMeans(df2[, c("T_FHC_Max1", "T_FHC_Max2", "T_FHC_Max3", "T_FHC_Max4")], na.rm = TRUE)
df2$T_Max_avg <- rowMeans(df2[, c("T_Max1", "T_Max2", "T_Max3", "T_Max4")], na.rm = TRUE)
df2$T_OR_avg <- rowMeans(df2[, c("T_OR1", "T_OR2", "T_OR3", "T_OR4")], na.rm = TRUE)
df2$T_OR_Max_avg <- rowMeans(df2[, c("T_OR_Max1", "T_OR_Max2", "T_OR_Max3", "T_OR_Max4")], na.rm = TRUE)
```


Average clean

```{r}
# For flir_group1_clean
write.csv(df, file = "/Users/karibautista/Desktop/df.csv", row.names = FALSE)
cat("\nDatasets have been saved as CSV files on the Desktop.\n")
```

```{r}
write.csv(df2, file = "/Users/karibautista/Desktop/df2.csv", row.names = FALSE)
```



# Combine df and df2 into a single dataframe
```{r}
combined_df <- rbind(df, df2)
```

# Save the combined dataframe to a CSV file
```{r}
write.csv(combined_df, "/Users/karibautista/Desktop/combined_data.csv", row.names = FALSE)
```




#----------------------------------------------------------------------------------------------------



# LONG TABLE TRANSFORMATION

## Transform datasets into long format
```{r}
library(tidyr)
transform_to_long <- function(data, base_vars, subject_vars) {
  
  long_data <- pivot_longer(
    data, 
    cols = starts_with(base_vars),      # Columns to gather (all temperature variables)
    names_to = c(".value", "Round"),    # ".value" means the base variable name stays as is, "Round" becomes the round number
    names_pattern = "(.*)([1-4])"       # Extract the base variable and round number
  )
  
  long_data$Round <- factor(long_data$Round, levels = c("1", "2", "3", "4"))
  
  return(long_data)
}
```


# Base temperature variables
```{r}
base_temp_vars <- c("T_offset", "Max1R13_", "Max1L13_", "aveAllR13_", "aveAllL13_", 
                    "T_RC", "T_RC_Dry", "T_RC_Wet", "T_RC_Max", 
                    "T_LC", "T_LC_Dry", "T_LC_Wet", "T_LC_Max", 
                    "RCC", "LCC", "canthiMax", "canthi4Max", 
                    "T_FHCC", "T_FHRC", "T_FHLC", "T_FHBC", "T_FHTC", 
                    "T_FH_Max", "T_FHC_Max", "T_Max", "T_OR", "T_OR_Max")

subject_vars <- c("SubjectID", "Gender", "Age", "Ethnicity", "Cosmetics", "T_atm", "Humidity", "Distance", "Date", "Time")
```


# Apply the function
```{r}
flir_group1_long <- transform_to_long(flir_group1_clean, base_temp_vars, subject_vars)
flir_group2_long <- transform_to_long(flir_group2_clean, base_temp_vars, subject_vars)
ici_group1_long  <- transform_to_long(ici_group1_clean, base_temp_vars, subject_vars)
ici_group2_long  <- transform_to_long(ici_group2_clean, base_temp_vars, subject_vars)

head(flir_group1_long)
head(flir_group2_long)
head(ici_group1_long)
head(ici_group2_long)

summary(flir_group1_long)
summary(flir_group2_long)
summary(ici_group1_long)
summary(ici_group2_long)

dim(flir_group1_long)
dim(flir_group2_long)
dim(ici_group1_long)
dim(ici_group2_long)

colnames(flir_group1_long)
colnames(flir_group2_long)
colnames(ici_group1_long)
colnames(ici_group2_long)

sum(is.na(flir_group1_long))
sum(is.na(flir_group2_long))
sum(is.na(ici_group1_long))
sum(is.na(ici_group2_long))
```

# Save cleaned datasets to CSV files

```{r}
# For flir_group1_clean
write.csv(flir_group1_long, file = "/Users/karibautista/Desktop/flir_group1_clean.csv", row.names = FALSE)

# For flir_group2_clean
write.csv(flir_group2_long, file = "/Users/karibautista/Desktop/flir_group2_clean.csv", row.names = FALSE)

# For ici_group1_clean
write.csv(ici_group1_long, file = "/Users/karibautista/Desktop/ici_group1_clean.csv", row.names = FALSE)

# For ici_group2_clean
write.csv(ici_group2_long, file = "/Users/karibautista/Desktop/ici_group2_clean.csv", row.names = FALSE)


cat("\nDatasets have been saved as CSV files on the Desktop.\n")
```






# Plot
#---------------#
# FLIR Group 1: 
#---------------#

# Set up a grid of plots
# y = 
#par(mfrow = c(2, 2))  # 1 row, 3 columns; adjust as necessary

# Example scatter plots
#plot(flir_group2_clean$Max1R13_1, flir_group2_clean$aveOralM, main = "Scatter Plot of Y vs. X1", xlab = "X1", ylab = "Y")
#plot(flir_group2_clean$Max1L13_1, flir_group2_clean$aveOralM, main = "Scatter Plot of Y vs. X2", xlab = "X2", ylab = "Y")
#plot(flir_group2_clean$aveAllR13_1, flir_group2_clean$aveOralM, main = "Scatter Plot of Y vs. X3", xlab = "X3", ylab = "Y")

# Reset the layout
#par(mfrow = c(1, 1))






