---
title: "Stepwise Regression DANA4820 Project"
author: "Carlos"
date: "2024-11-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Define the file path

train_data <- read.csv("train_data.csv")
View(train_data)
```

## Rename the variables
```{r}
# Rename variables in the dataset
colnames(train_data) <- sub("T_OR_Max_avg", "Max_t_mouth", colnames(train_data))
colnames(train_data) <- sub("T_FHC_Max_avg", "Max_t_ext_fh", colnames(train_data))
colnames(train_data) <- sub("T_FHBC_avg", "Ave_t_bot_fh", colnames(train_data))
colnames(train_data) <- sub("T_FHTC_avg", "Ave_t_top_fh", colnames(train_data))
colnames(train_data) <- sub("RCC_avg", "Ave_t_r_inner_canthus", colnames(train_data))
colnames(train_data) <- sub("aveAllR_avg", "Ave_t_r_canthus", colnames(train_data))
colnames(train_data) <- sub("aveAllL_avg", "Ave_t_l_canthus", colnames(train_data))
colnames(train_data) <- sub("T_LC_Wet_avg", "Ave_t_l_canthus_w", colnames(train_data))
colnames(train_data) <- sub("canthi4Max_avg", "Max_t_ext_canthus", colnames(train_data))
colnames(train_data) <- sub("T_FHRC_avg", "Ave_r_left_fh", colnames(train_data))
colnames(train_data) <- sub("T_FHLC_avg", "Ave_t_left_fh", colnames(train_data))
colnames(train_data) <- sub("T_Max_avg", "Max_t_face", colnames(train_data))
colnames(train_data) <- sub("Gender", "gender", colnames(train_data))
colnames(train_data) <- sub("Age", "age", colnames(train_data))
colnames(train_data) <- sub("Ethnicity", "ethnic", colnames(train_data))
colnames(train_data) <- sub("Cosmetics", "cosm", colnames(train_data))
colnames(train_data) <- sub("T_atm", "Ambient_t", colnames(train_data))
colnames(train_data) <- sub("Humidity", "humidity", colnames(train_data))
```


# 1. Models types
```{r}
# Fit the null model (with no predictors)
null_model <- lm(aveOralM ~ 1, data = train_data)

# Fit the full model (with all explanatory variables)
full_model <- lm(
  aveOralM ~ Max_t_mouth + Max_t_ext_fh + Ave_t_bot_fh + Ave_t_top_fh +
    Ave_t_r_inner_canthus + Ave_t_r_canthus + Ave_t_l_canthus + Ave_t_l_canthus_w +
    Max_t_ext_canthus + Ave_r_left_fh + Ave_t_left_fh + Max_t_face +
    gender + age + ethnic + cosm + Ambient_t + humidity,
  data = train_data
)

# Forward stepwise regression
forward_model <- step(
  null_model,
  scope = list(lower = ~1, upper = formula(full_model)),
  direction = "forward",
  trace = FALSE
)

# Backward stepwise regression
backward_model <- step(
  full_model,
  direction = "backward",
  trace = FALSE
)

# Stepwise regression in both directions
both_model <- step(
  null_model,
  scope = list(lower = ~1, upper = formula(full_model)),
  direction = "both",
  trace = FALSE
)

# Print model summaries with interpretations
cat("\n### Summary of Forward Model ###\n")
summary(forward_model)
cat("2. Final Model:\n")
cat(deparse(formula(forward_model)), "\n")

cat("\n### Summary of Backward Model ###\n")
summary(backward_model)
cat("2. Final Model:\n")
cat(deparse(formula(backward_model)), "\n")

cat("\n### Summary of Both Directions Model ###\n")
summary(both_model)
cat("2. Final Model:\n")
cat(deparse(formula(both_model)), "\n")

cat("\n### Summary of Full Model ###\n")
summary(full_model)
cat("2. Final Model:\n")
cat(deparse(formula(full_model)), "\n")

# Print AIC values for comparison
aic_forward <- AIC(forward_model)
aic_backward <- AIC(backward_model)
aic_both <- AIC(both_model)
aic_full <- AIC(full_model)

cat("\n### AIC Values for Model Comparison ###\n")
cat("Forward Model AIC:", aic_forward, "\n")
cat("Backward Model AIC:", aic_backward, "\n")
cat("Both Directions Model AIC:", aic_both, "\n")
cat("Full Model AIC:", aic_full, "\n")
```
# 2. Multicollinearity
```{r}
library(car)
vif(full_model)
vif(forward_model)
vif(backward_model)
vif(both_model)
```


# 3. Anova with the 4 models
```{r}
anova(full_model,forward_model)
anova(full_model,backward_model)
anova(full_model,both_model)

```
*


# 4. Residual Plots

#### - Full Model
```{r}
#Get residuals and predicted values
residuals1 <- resid(full_model)  # Get residuals
predicted_values1 <- predict(full_model)  # Get predicted values

# Calculate the standard deviation of the residuals
std_dev1 <- sd(residuals1)

# Step 1: Identify outliers that lie above 2 standard deviations or below -2 standard deviations
outliers1 <- residuals1 > (2 * std_dev1) | residuals1 < (-2 * std_dev1)

# Step 2: Count the number of outliers
num_outliers1 <- sum(outliers1)

# Step 3: Calculate the percentage of outliers
percentage_outliers1 <- (num_outliers1 / length(residuals1)) * 100

# Step 4: Output the percentage of outliers
cat("Percentage of residuals that are outliers (beyond ±2 SD):", percentage_outliers1, "%\n")

# Step 5: Check if the percentage of outliers is greater than 5%
if (percentage_outliers1 > 5) {
  cat("More than 5% of the residuals are outliers.\n")
} else {
  cat("Less than or equal to 5% of the residuals are outliers.\n")}

# Step 6: Create the residual plot
plot(predicted_values1, residuals1, 
     ylab = "Residuals", 
     xlab = "Fitted Values (Oral M Temp)", 
     main = "Residual Full Model Plot")

# Add a horizontal line at 0
abline(h = 0, col = "red", lty = 2)

# Add lines for ±2 standard deviations
abline(h = 2 * std_dev1, col = "red", lty = 2)   # Upper line
abline(h = -2 * std_dev1, col = "red", lty = 2)   # Lower line

```
95% of the dots are within the two standard deviations

#### - Forward (winner)
```{r}
#Get residuals and predicted values
residuals2 <- resid(forward_model)  # Get residuals
predicted_values2 <- predict(forward_model)  # Get predicted values

std_dev2 <- sd(residuals2)
outliers2 <- residuals2 > (2 * std_dev2) | residuals2 < (-2 * std_dev2)
num_outliers2 <- sum(outliers2)
percentage_outliers2 <- (num_outliers2 / length(residuals2)) * 100

cat("Percentage of residuals that are outliers (beyond ±2 SD):", percentage_outliers2, "%\n")

if (percentage_outliers2 > 5) {
  cat("More than 5% of the residuals are outliers.\n")
} else {
  cat("Less than or equal to 5% of the residuals are outliers.\n")}

# Step 6: Create the residual plot
plot(predicted_values2, residuals2, 
     ylab = "Residuals", 
     xlab = "Fitted Values (Oral M Temp)", 
     main = "Residual Forward Plot")

# Add a horizontal line at 0
abline(h = 0, col = "red", lty = 2)

# Add lines for ±2 standard deviations
abline(h = 2 * std_dev2, col = "red", lty = 2)   # Upper line
abline(h = -2 * std_dev2, col = "red", lty = 2)   # Lower line

```
```{r}
plot(forward_model)
```


#### - Backward Model
```{r}
#Get residuals and predicted values
residuals3 <- resid(backward_model)  # Get residuals
predicted_values3 <- predict(backward_model)  # Get predicted values

std_dev3 <- sd(residuals3)
outliers3 <- residuals3 > (2 * std_dev3) | residuals3 < (-2 * std_dev3)
num_outliers3 <- sum(outliers3)
percentage_outliers3 <- (num_outliers3 / length(residuals3)) * 100

cat("Percentage of residuals that are outliers (beyond ±2 SD):", percentage_outliers3, "%\n")

if (percentage_outliers3 > 5) {
  cat("More than 5% of the residuals are outliers.\n")
} else {
  cat("Less than or equal to 5% of the residuals are outliers.\n")}

# Step 6: Create the residual plot
plot(predicted_values3, residuals3, 
     ylab = "Residuals", 
     xlab = "Fitted Values (Oral M Temp)", 
     main = "Residual Backward Plot")

# Add a horizontal line at 0
abline(h = 0, col = "red", lty = 2)

# Add lines for ±2 standard deviations
abline(h = 2 * std_dev3, col = "red", lty = 2)   # Upper line
abline(h = -2 * std_dev3, col = "red", lty = 2)   # Lower line
```

#### - Both model
```{r}
#Get residuals and predicted values
residuals4 <- resid(both_model)  # Get residuals
predicted_values4 <- predict(both_model)  # Get predicted values

std_dev4 <- sd(residuals4)
outliers4 <- residuals4 > (2 * std_dev4) | residuals4 < (-2 * std_dev4)
num_outliers4 <- sum(outliers4)
percentage_outliers4 <- (num_outliers4 / length(residuals4)) * 100

cat("Percentage of residuals that are outliers (beyond ±2 SD):", percentage_outliers4, "%\n")

if (percentage_outliers4 > 5) {
  cat("More than 5% of the residuals are outliers.\n")
} else {
  cat("Less than or equal to 5% of the residuals are outliers.\n")}

# Step 6: Create the residual plot
plot(predicted_values4, residuals4, 
     ylab = "Residuals", 
     xlab = "Fitted Values (Oral M Temp)", 
     main = "Residual Both Plot")

# Add a horizontal line at 0
abline(h = 0, col = "red", lty = 2)

# Add lines for ±2 standard deviations
abline(h = 2 * std_dev4, col = "red", lty = 2)   # Upper line
abline(h = -2 * std_dev4, col = "red", lty = 2)   # Lower line
```


# WINNER MODEL

lm(formula = aveOralM ~ Max_t_face + Ambient_t + Ave_t_l_canthus_w, 
    data = train_data)
  
# 5. Transformation  
    
###  a) Box-Cox
Find the best transformation

Based on this, Box Cox suggests to make a log transformation

###### - Transformation
```{r}
library(MASS)

# Apply Box-Cox to the model
boxcox_result <- boxcox(forward_model, plotit = TRUE)  # plotit = FALSE suppresses the plot

# Find the optimal lambda (the value of lambda that maximizes the log-likelihood)
lm_best <- boxcox_result$x[which.max(boxcox_result$y)]
lm_best

train_data$new_var <- ifelse(lm_best == 0,
                              log(train_data$aveOralM), 
                              (train_data$aveOralM^lm_best - 1) /lm_best)


model_trans = lm(new_var ~ Max_t_face + Ambient_t + Ave_t_l_canthus_w, 
                 data = train_data)

summary(model_trans)
```
### - Assumption for Box Cox
```{r}
plot(model_trans)
```
```{r}
#Get residuals and predicted values
residuals5 <- resid(model_trans)  # Get residuals
predicted_values5 <- predict(model_trans)  # Get predicted values

std_dev5 <- sd(residuals5)
outliers5 <- residuals5 > (2 * std_dev5) | residuals5 < (-2 * std_dev5)
num_outliers5 <- sum(outliers5)
percentage_outliers5 <- (num_outliers5 / length(residuals5)) * 100

cat("Percentage of residuals that are outliers (beyond ±2 SD):", percentage_outliers5, "%\n")

if (percentage_outliers5 > 5) {
  cat("More than 5% of the residuals are outliers.\n")
} else {
  cat("Less than or equal to 5% of the residuals are outliers.\n")}

# Step 6: Create the residual plot
plot(predicted_values5, residuals5, 
     ylab = "Residuals", 
     xlab = "Fitted Values (Oral M Temp)", 
     main = "Residual Plot")

# Add a horizontal line at 0
abline(h = 0, col = "red", lty = 2)

# Add lines for ±2 standard deviations
abline(h = 2 * std_dev5, col = "red", lty = 2)   # Upper line
abline(h = -2 * std_dev5, col = "red", lty = 2)   # Lower line
```


### b) Log

```{r}
lm_log <- lm(log(aveOralM) ~ Max_t_face + Ambient_t + Ave_t_l_canthus_w, 
             data = train_data)
summary(lm_log)
```
```{r}
AIC(lm_log)
```

**NOTE: we cannot perform anova, since the y's do not have the same values or compare AICs




